MinCall --- MinION end2end deep learning basecaller

### 2017-06-30 paper/main.tex
Edited introduction/sequecing overview

Although MinION is able to produce long reads~\cite{loman1-100k,loman2-800k}, they have a high sequencing error rate. This has been somewhat alleviated with new R9 pore model, replacing older R7 ones. In this paper, we show that this error rate can be reduced by replacing the default base caller provided by the manufacturer with a properly trained neural network model. In the future new R9.5 chemistry and 1D\^2 reads should supersede current models.

In the MinION device, single-stranded DNA fragments move through nanopores, which causes drops in the electric current. The electric current is measured at each pore several thousand times per second, 4000 times exactly in our dataset. The electric current depends mostly on the context of several DNA bases passing through the pore at the time of measurement. As the DNA moves through the pore, the context shifts, and the electric current changes.

The MinION device typically yields reads several thousands bases long, even couple hundred thousand bases long reads were repoted~\cite{loman1-100k,loman2-800k}. However the cost in on accuracy, signicifantly lower than older, more reliable and expensive sequencing methods.

@@

Electrical current runs through the nanopore and exact nucleotides context within influences the nanopore's resistance. This resistive effect is our sensor data, that is the current fluctuations as DNA passes though the pore. The nanopore is 6 nucleotides wide, and many models use this information in their approaches, while we're opted out of this technicality.

@@

\section{Basecalling}

The core of the decoding process is the basecalling step. Nowdays there's multiple basecalling options, what official and unoffical ones.

>>>>

Although MinION is able to produce long reads, even up to 882 kb~\cite{loman1-100k,loman2-800k}, they have a high sequencing error rate. This has been somewhat alleviated with new R9 pore model, replacing older R7 ones. In this paper, we show that this error rate can be reduced by our approach with properly trained neural network model.

@@

As DNA strand pass through the nanopore, they are propelled by the current. However this current varies depending on specific DNA nucletide context within the nanopore, changing its resistance. We sample the curreny multiple times per second, 4000Hz in our dataset, and from this data we deduce passing DNA fragement. By design, nanopore is 6 nucleotides wide, and many models use this information internally, yet we opted our of it in hopes of simpler model, less feature engineering and better homopolymer detection.

@@

\section{Basecalling}
The core of the decoding process is the basecalling step, that is translating the current sample to passed nucleotided sequence. Nowdays there's multiple basecalling options, what official and unoffical ones.
